---
title: "LBB"

output: html_notebook
---

Inspiration
The data-set aims to answer the following key questions:

1. Does various predicting factors which has been chosen initially really affect the Life expectancy? What are the predicting variables actually affecting the life expectancy?
2. Should a country having a lower life expectancy value(<65) increase its healthcare expenditure in order to improve its average lifespan?
3. Does Life Expectancy has positive or negative correlation with eating habits, lifestyle, exercise, smoking, drinking alcohol etc.
4. Do densely populated countries tend to have lower life expectancy?

```{r}
library(dplyr)
library(ggplot2)
library(GGally)
library(tidyr)
library(MLmetrics)
library(car)
```

```{r}
options(scipen = 999)
```


```{r}
life <- read.csv("Life Expectancy Data.csv")
```
```{r}
life <- life %>% 
  drop_na() %>% 
  mutate(Country = as.factor(Country),
         Status = as.factor(Status),
         Year = as.factor(Year))
```

```{r}
str(life)
```

```{r}
colSums(is.na(life))
```
```{r}
life_low <- life %>% 
  filter(Life.expectancy < 65) %>% 
  select(c("Life.expectancy","Total.expenditure"))
```
```{r}
model_low <- lm(formula = Life.expectancy~Total.expenditure, data = life_low)
summary(model_low)
```

```{r}
plot(life_low)
abline(model_low)
```
```{r}
cor(life_low)
```
> negative correlation, so there is no need to increase expenditure

```{r}
life_exp <- life %>% 
  select(-c("Country","Year","Status"))
```


```{r}
str(life_exp)
```
```{r}
sapply(life, n_distinct)
```


```{r warning=FALSE}
ggplot(data=life,mapping=aes(Year,Life.expectancy,color=Status))+
  geom_point()+
  scale_color_brewer(palette="Accent")+
  geom_smooth(method="lm",se=FALSE)+
  labs(title="Life Expectancy from 2000-2015")+
  theme(plot.title = element_text(hjust = 0.5))
```
```{r}
ggplot(data=life,mapping=aes(Year,Adult.Mortality,color=Status))+
  geom_point()+
  scale_color_brewer(palette="Accent")+
  geom_smooth(method="lm",se=FALSE)+
  labs(title="Adult Mortality from 2000-2015")+
  theme(plot.title = element_text(hjust = 0.5))
```
```{r}
ggcorr(life_exp, label = TRUE, label_size = 3, hjust = 1, size = 3, color = "grey50")
```
>> Kesimpulan :
Schooling , Income Composition of Resources, Adult.Mortality, HIV.AIDS, BMI


```{r}
model_all <- lm(formula = Life.expectancy~., data = life_exp)
summary(model_all)
```
```{r}
vif(model_all)
```
```{r}
life_exp <- life_exp %>% 
  select(-c("percentage.expenditure","under.five.deaths"))
```

```{r}
set.seed(123)
samplesize <- round(0.7 * nrow(life_exp), 0)
index <- sample(seq_len(nrow(life_exp)), size = samplesize)

data_train <- life_exp[index, ]
data_test <- life_exp[-index, ]
```

```{r}
model_all1 <- lm(formula = Life.expectancy~., data = data_train)
summary(model_all1)
```
```{r}
vif(model_all1)

```

```{r}
model_cor <-  lm(formula = Life.expectancy~Schooling + Income.composition.of.resources + HIV.AIDS + BMI + Adult.Mortality, data = data_train)
summary(model_cor)
```
```{r}
model_dis <- lm(formula = Life.expectancy~ Alcohol + Hepatitis.B + Measles + 
    Polio + Diphtheria + HIV.AIDS , data = data_train)
summary(model_dis)
```

```{r}
summary(model_all)$adj.r.squared
summary(model_cor)$adj.r.squared
summary(model_dis)$adj.r.squared
```
> Kesimpulan: Pada kasus ini, model dengan seluruh kolom sebagai prediktor adalah model terbaik

```{r}
data_train$Prediksi <- predict(model_all1, newdata = data_train %>% select(-Life.expectancy))
```

```{r}
res_train <- data_train$Prediksi - data_train$Life.expectancy
data.frame(prediksi = data_train$Prediksi,
           actual = data_train$Life.expectancy,
           residual = res_train)
```

```{r}
RMSE(model_all1$fitted.values, data_train$Life.expectancy)
```
```{r}
MAPE(data_train$Prediksi, data_train$Life.expectancy) * 100
```


```{r}
data_test$Prediksi <- predict(model_all1, newdata = data_test %>% select(-Life.expectancy))
```

```{r}
res_test <- data_test$Prediksi - data_test$Life.expectancy
w <- data.frame(prediksi = data_test$Prediksi,
           actual = data_test$Life.expectancy,
           residual = res_test)
```

```{r warning=FALSE}
RMSE(model_all1$fitted.values, data_test$Life.expectancy)
```

```{r}
MAPE(data_test$Prediksi, data_test$Life.expectancy) * 100
```
```{r}
set.seed(123)
samplesize <- round(0.7 * nrow(life_exp), 0)
index <- sample(seq_len(nrow(life_exp)), size = samplesize)

data_train1 <- life_exp[index, ]
data_test1 <- life_exp[-index, ]
```

```{r}
# model tanpa prediktor
model_life_non <- lm(formula = Life.expectancy ~ 1, data = data_train1)
# model dengan semua prediktor
model_life_all <- lm(formula = Life.expectancy ~ ., data = data_train1)
```

```{r}
both_model <- step(model_life_non, direction = "both",
     scope = list(lower = model_life_non, upper = model_life_all), trace = 0)

summary(both_model)
```

```{r}
# Backward elimination
backward_model <- step(model_life_all, trace = 0)
summary(backward_model)
```

```{r}

forward_model <- step(model_life_non, direction = "forward",
     scope = list(lower = model_life_non, upper = model_life_all), trace = 0)
summary(forward_model)
```

```{r}
library(performance)
compare_performance(backward_model, forward_model, both_model)
```
> all model is the same

```{r}
data_train1$Prediksi <- predict(both_model, newdata = data_train1 %>% select(-Life.expectancy))
```

```{r}
res_train1 <- data_train1$Prediksi - data_train1$Life.expectancy
data.frame(prediksi = data_train1$Prediksi,
           actual = data_train1$Life.expectancy,
           residual = res_train1)
```

```{r}
RMSE(both_model$fitted.values, data_train1$Life.expectancy)
```
```{r}
MAPE(data_train1$Prediksi, data_train1$Life.expectancy) * 100
```
```{r}
data_test1$Prediksi <- predict(both_model, newdata = data_test1 %>% select(-Life.expectancy))
```

```{r}
res_test1 <- data_test1$Prediksi - data_test1$Life.expectancy
h <- data.frame(prediksi = data_test1$Prediksi,
           actual = data_test1$Life.expectancy,
           residual = res_test1)
```

```{r warning=FALSE}
RMSE(both_model$fitted.values, data_test1$Life.expectancy)
```

```{r}
MAPE(data_test1$Prediksi, data_test1$Life.expectancy) * 100
```

```{r}
hist(h$residual, breaks=30)
```
>normal distribution

H0: error berdistribusi normal
H1: error tidak berdistribusi normal
(harapannya pvalue > alpha agar keputusan yang diambil adalah gagal tolak H0)

```{r}
shapiro.test(h$residual)
```
> p value < 0.005
> tolak H0

```{r warning=FALSE}
library(lmtest)
plot(data_test1$Life.expectancy, h$residual)
abline(h = 0, col = "red")
```
$$
H_0: Variansi\ error\ menyebar\ konstan\ (Homoscedasticity)\\
H_1: Variansi\ error\ menyebar\ tidak\ konstan\ membentuk\ pola\ (Heteroscedasticity)
$$
Breusch-Pagan hypothesis test: (harapannya p-value > alpha agar gagal tolak H0)

```{r}
bptest(both_model)
```
```{r}
vif(both_model)
```







